---
title: "Lab3_KatieTumang"
author: "Katie Tumang"
date: "September 7, 2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

```{r}
download.file(url = "https://ndownloader.figshare.com/files/2292169",
              destfile = "data/portal_data_joined.csv")
```

```{r}
surveys <- read.csv("data/portal_data_joined.csv")
head(surveys)
View(surveys)
str(surveys)
```

## Challenge
### What is the class of the object surveys? Data frame
### How many rows and how many columns are in this object? 34786 rows, 13 columns
### How many species have been recorded during these surveys? 48 species

## Indexing and subsetting data frames

```{r}
surveys[1:3, 1] 
```

```{r}
head_surveys <- surveys [1:6, ]
head_surveys
```

```{r}
head_surveys[, -1]
```

```{r}
tail(surveys)[,"species_id"]
```

## Challenge 2
### 1.
```{r}
surveys_200 <- surveys [200, ]
```

#### 2.
```{r}
n_rows <- nrow(surveys)
surveys_last <- surveys[n_rows, ]
surveys_last
```

```{r}
tail(surveys)
```

#### 3.
```{r}
surveys_middle <- surveys[n_rows/2, ]
surveys_middle
```

#### 4.
```{r}
surveys_head <- surveys[-(7:n_rows),]
surveys_head
```

```{r}
head(surveys)
```

## Factors

```{r}
sex <- factor(c("male", "female", "female", "male"))
levels(sex)
nlevels(sex)
```

```{r}
sex # current order
```

```{r}
sex <- factor(sex, levels = c("male", "female"))
sex # after re-ordering
```

#### Converting factors

```{r}
as.character(sex)
```

```{r}
year_fct <- factor(c(1990, 1983, 1977, 1998, 1990))
as.numeric(levels(year_fct))[year_fct]
```

#### Renaming factors

```{r}
## bar plot of the number of females and males captured during the experiment:
plot(as.factor(surveys$sex))
```

```{r}
sex <- factor(surveys$sex)
head(sex)
levels(sex)
levels(sex)[1] <- "undetermined"
levels(sex)
head(sex)
```

### Challenge 3.
#### 1. Rename "F" and "M" to "female" and "male" respectively
```{r}
levels(sex)[2:3] <- c("female","male")
head(sex)
```
#### 2. Make a bar plot where "undetermined" is last
```{r}
sex <- factor(sex, levels = c("female","male","undetermined"))
plot(sex)
```

### Using stringAsFactors = FALSE

```{r}
surveys <- read.csv("data/portal_data_joined.csv", stringsAsFactors = TRUE)
str(surveys)
surveys <- read.csv("data/portal_data_joined.csv", stringsAsFactors = FALSE)
str(surveys)
surveys$plot_type <- factor(surveys$plot_type)
```

### Challenge 4.
#### 1.

```{r}
animal_data <- data.frame(
      animal = c("dog", "cat", "sea cucumber", "sea urchin"),
      feel = c("furry", "furry", "squishy", "spiny"),
      weight = c(45, 8, 1.1, 0.8)
)
```

#### 2.
```{r}
country_climate <- data.frame(
       country = c("Canada", "Panama", "South Africa", "Australia"),
       climate = c("cold", "hot", "temperate", "hot/temperate"),
       temperature = c(10, 30, 18, 15),
       northern_hemisphere = c(TRUE, TRUE, FALSE, FALSE),
       has_kangaroo = c(FALSE, FALSE, FALSE, TRUE)
       )
```

```{r}
str(country_climate)
```

##### 2a. I expected country, climate, northern_hemisphere, and has_kangaroo to be factors, and temperature to be numeric; country, climate, temperature, and northern_hemisphere were factors, while has_kangaroo was numeric. My initial assumption was wrong, as I forgot that despite being a numeric value, temperature is a factor that can be ordered and represents more than a number. Likewise, northern_hemisphere is a factor that categorizes whether a country is in the northern hemisphere, while has_kangaroo is a yes or no question, not a factor.
##### 2b. Adding stringsAsFactors = FALSE would have produced character vectors instead of factors.
##### 2c. To ensure that each column has the accurate data type, one would remove the quotations from FALSE in northern_hemisphere, and replace 1 with TRUE in has_kangaroo.

### Formatting Dates
```{r}
str(surveys)
```

```{r}
library("lubridate")
```

```{r}
my_date <- ymd("2020-01-01")
str(my_date)
```

```{r}
my_date <- ymd(paste("2020", "1", "1", sep = "-"))
str(my_date)
```

```{r}
paste(surveys$year, surveys$month, surveys$day, sep = "-")
ymd(paste(surveys$year, surveys$month, surveys$day, sep = "-"))
surveys$date <- ymd(paste(surveys$year, surveys$month, surveys$day, sep = "-"))
str(surveys)
summary(surveys$date)
```

```{r}
missing_dates <- surveys[is.na(surveys$date),c("year","month","day")]
head(missing_dates)
```

