---
title: "Lab3_KatieTumang_3b"
author: "Katie Tumang"
date: "September 7, 2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

```{r}
surveys <- read_csv("data/portal_data_joined.csv")
```

### Inspect the data
```{r}
str(surveys)
```

### Preview the data
```{r}
View(surveys)
```

```{r}
select(surveys, plot_id, -species_id, weight)
filter(surveys, year == 1995)
```

### Intermediate steps
```{r}
surveys2 <- filter(surveys, weight < 5)
surveys_sml <- select(surveys2, species_id, sex, weight)
```

### Nested functions
```{r}
surveys_sml <- select(filter(surveys, weight < 5), species_id, sex, weight)
```

### Pipes
```{r}
surveys_sml <- surveys %>% 
  filter(weight < 5) %>% 
  select(species_id, sex, weight)

surveys_sml
```

### Challenge
```{r}
surveys %>% 
  filter(year < 1995) %>% 
  select(year, sex, weight)
```

### Mutate
```{r}
surveys %>%
  filter(!is.na(weight)) %>% 
  mutate(weight_kg = weight / 1000,
  weight_lb = weight_kg * 2.2) %>% 
  head()
```

### Challenge 2
```{r}
surveys_hindfoot_cm <- surveys %>% 
  filter(!is.na(hindfoot_length)) %>% 
  mutate(hindfoot_cm = hindfoot_length / 10) %>% 
  filter(hindfoot_cm < 3) %>% 
  select(species_id, hindfoot_cm)
surveys_hindfoot_cm
```

### Split-apply-combine data analysis
```{r}
surveys %>%
  filter(!is.na(weight)) %>% 
  group_by(sex, species_id) %>% 
  summarize(mean_weight = mean(weight),
            min_weight = min(weight)) %>% 
  arrange(desc(mean_weight))
```

### Counting
```{r}
surveys %>%
  count(sex, species) %>% 
  arrange(species, desc(n))
```

### Challenge 3
#### 1. How many animals were caught in each plot_type surveyed?
```{r}
surveys %>% 
  count(plot_type)
```

#### 2. Mean, min, max hindfoot length, # of observations:
```{r}
surveys %>%
  filter(!is.na(hindfoot_length)) %>% 
  group_by(species_id) %>% 
  summarize(
    mean_hindfoot_length = mean(hindfoot_length),
    max_hindfoot_length = max(hindfoot_length),
    min_mindfoot_length = min(hindfoot_length),
    n = n()
  )
```

#### 3. What was the heaviest animal measured in each year?
```{r}
surveys %>% 
  filter(!is.na(weight)) %>% 
  group_by(year) %>% 
  filter(weight == max(weight)) %>% 
  select(year, genus, species_id, weight) %>% 
  arrange(year)
```

### Reshaping with gather and spread
```{r}
surveys_gw <- surveys %>% 
  filter(!is.na(weight)) %>%
  group_by(plot_id, genus) %>%
  summarize(mean_weight = mean(weight))
str(surveys_gw)
```

```{r}
surveys_spread <- surveys_gw %>% 
  spread(key = genus, value = mean_weight)
str(surveys_gw)
```

```{r}
surveys_gw %>% 
  spread(genus, mean_weight, fill = 0) %>% 
  head()
```

### Gathering
```{r}
surveys_gather <- surveys_spread %>% 
  gather(key = "genus", value = "mean_weight", -plot_id)
str(surveys_gather)
```

### Challenge 4
#### 1.
```{r}
surveys_genera <- surveys %>% 
  group_by(plot_id, year) %>% 
  summarize(n_genera = n_distinct(genus)) %>% 
  spread(year, n_genera)
head(surveys_genera)
```

#### 2.
```{r}
surveys_genera %>% 
  gather("year","n_genera", -plot_id)
str(surveys_genera)
```

#### 3.
```{r}
surveys_measure <- surveys %>% 
  gather("measurement", "value", hindfoot_length, weight)
str(surveys_measure)
```

#### 4.
```{r}
surveys_measure %>%
  group_by(year, measurement, plot_type) %>% 
  summarize(mean_value = mean(value, na.rm = TRUE)) %>% 
  spread(measurement, mean_value)
str(surveys_measure)
```

### Exporting data
```{r}
surveys_complete <- surveys %>%
  filter(!is.na(weight),
         !is.na(hindfoot_length),
         !is.na(sex))
```

```{r}
species_count <- surveys_complete %>%
  count(species_id) %>% 
  filter(n >= 50)
```


```{r}
surveys_complete <- surveys_complete %>% 
  filter(species_id %in% species_count$species_id)
dim(surveys_complete)

write_csv(surveys_complete, path = "data/surveys_complete.csv")
```

