---
title: "Lab8_KatieTumang"
author: "Katie Tumang"
date: "November 28, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Use NGPhylogeny.fr to analysis the set of rRNA sequence provided. Describe the methods and put the .png file from your analysis into your Lab 8 .Rmd file

## ggtree tutorial

```{r}
library(tidyverse)
library(ggtree)
library(treeio)
library(ggimage)
```

```{r}
install.packages("treeio", dependencies = TRUE, INSTALL_opts = '--no-lock')
library(treeio)
```


```{r}
library(tidyverse)
library(ggtree)
tree <- read.tree("tree_newick.nwk")
tree
```

```{r}
ggtree(tree, color="grey", size=1, linetype=1) + theme_tree2()

```

### Exercise 1

#### 1. Create a slanted phylogenetic tree.

```{r}
ggtree(tree, color="purple", layout="slanted") 
```

#### 2. Create a circular phylogenetic tree.

```{r}
ggtree(tree, color="purple", layout="circular")
```

#### 3. Create a circular unscaled cladogram with thick red lines.

```{r}
ggtree(tree, color="red", branch.length="none", size=3, layout="circular")
```

```{r}
p <- ggtree(tree)
p + geom_nodepoint()
p + geom_tippoint()
p + geom_tiplab()
```

#### Exercise 2

```{r}
p <- ggtree(tree)
p + geom_nodepoint(size=5, color="yellow", alpha=0.5) + 
  geom_tippoint(color="purple", pch=18) + geom_tiplab(color="purple") +ggtitle("Exercise 2 tree")
```

### Tree annotation

```{r}
ggtree(tree) + geom_text(aes(label=node), hjust=-.3)
```

```{r}
ggtree(tree) + geom_tiplab()
MRCA(tree, c("C", "E"))
MRCA(tree, c("G", "H"))
```

### Labeling clades

```{r}
ggtree(tree) + 
  geom_tiplab() + 
  geom_cladelabel(node=17, label="Random clade", 
                  color="red2", offset=.8, align=TRUE) + 
  geom_cladelabel(node=21, label="Other clade", 
                  color="green2", offset=.8, align=TRUE) + 
  theme_tree2() + 
  xlim(0, 70) + 
  theme_tree()
```

```{r}
ggtree(tree) + 
  geom_tiplab() + 
  geom_hilight(node=17, fill="grey") + 
  geom_hilight(node=21, fill="purple")
```

### Connecting taxa

```{r}
ggtree(tree) + 
  geom_tiplab() + 
  geom_taxalink("E", "H", color="blue3") +
  geom_taxalink("C", "G", color="pink", curvature=-.9)
```

#### Exercise 3

##### 1. 
```{r}
MRCA(tree, c("B", "C"))
MRCA(tree, c("L", "J"))
```

```{r}
ggtree(tree) + 
  theme_tree2() +
  geom_tiplab() + 
  geom_hilight(node=19, fill="gold") + 
  geom_hilight(node=23, fill="purple") +
  geom_cladelabel(node=17, label="Superclade 17", 
                  color="red2", offset=.8, align=TRUE) + 
  geom_taxalink("C", "E", color="grey", linetype=2) +
  geom_taxalink("G", "J", color="grey", linetype=2, curvature=-.9)
  
```


### Advanced tree notation

```{r}
tree <- read.beast("flu_tree_beast.tree")
```

```{r}
ggtree(tree, mrsd="2013-01-01") + 
  theme_tree2() +
  geom_tiplab(align=TRUE, linesize=.5) + 
  xlim(1990, 2020)
```

```{r}
msaplot(p=ggtree(tree), fasta="flu_aasequence.fasta", window=c(150, 175)) + coord_polar(theta="y")
```

#### CIPRES tree

```{r}
tree <- read.tree("fastree_result.tre")
```

```{r}
ggtree(tree) +
  geom_tiplab(hjust=-.1, size = 3) +
  geom_hilight(node=34, fill="gold", extend = 1) + 
  geom_hilight(node=21, fill="purple", extend = 1)
```


