---
title: "Lab9_KatieTumang"
author: "Katie Tumang"
date: "11/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 14.3 Matching patterns with regular expressions

```{r}
library(stringr)
library(tidyverse)
```

### 14.3.1 basic matches

```{r}
x <- c("rabbit","parrot","frog")
str_view(x,"bb")
```

```{r}
str_view(x,".a.")
```

```{r}
period <- "\\."
writeLines(period)
```

```{r}
x <- c("abc","a.c","bef")
str_view(x,"a\\.c")
```

```{r}
x <- "b\\c"
writeLines(x)
str_view(x,"\\\\")
```

#### 14.3.1.1 exercises
##### 1. \ is used as an escape character. Escaping \ by writing \\ creates a regular expression. However, this needs a string, but the string needs to escape \ as well by using \\. Thus, to escape \ and create the expression using a string (itself escaping \), one would need to use 4 \'s: \\\\

### Anchors

```{r}
x <- c("rabbit","parrot","frog")
str_view(x,"^r")
```

```{r}
str_view(x,"t$")
```


```{r}
x <- c("rabbit pie","easter rabbit","rabbit")
str_view(x,"rabbit")
```

```{r}
str_view(x,"^apple$")
```

#### 14.3.2.1 Exercises
##### 1.
```{r}
x <- c("$^$","^^^")
str_view(x,"[$\\^$]")
```

### Character classes and alternatives

```{r}
x <- c("abc","a.c","a*c","a c")
str_view(x, "a[.]c")
```

```{r}
str_view(x,".[*]c")
```

```{r}
str_view(x,"a[ ]")
```

```{r}
x <- c("grey", "gray")
str_view(x, "gr(e|a)y")
```

#### 14.3.3.1 Exercises
##### 1.
```{r}
x <- c("california", "rabbit","aslkfhas","yanic","ggbgtr")
str_view(x, "[aeiou]")
```

```{r}
str_view(x, "[^aeiou]")
```

```{r}
x <- c("ted","peed","fed","feed")
str_view(x, "ed","^eed")
```

```{r}
x <- c("ted","rise","losing","surprise")
str_view(x, "ing|ise$")
```


### Repetition

```{r}
x <- "MDCCCLXXXVIII"
str_view(x,"CC?")
```

```{r}
str_view(x,"CC+")
```

```{r}
str_view(x,'C[LX]+')
```

```{r}
str_view(x, "C{2}")
```

```{r}
str_view(x, "C{2,}")
```

```{r}
str_view(x, "C{2,}")
```

```{r}
str_view(x, 'C{2,3}?')
```

```{r}
str_view(x, 'C[LX]+?')
```

##### 14.3.4.1 Exercises
###### 1. Describe the equivalents of ?, +, * in {m,n} form.
####### The character ? is equal to {,1}
####### The character + is equal to {1,}
####### The character * is equal to {0,}

### Grouping and backreferences

```{r}
str_view(fruit, "(..)\\1", match = TRUE)
```

### Tools

```{r}
x <- c("james", "buck", "charlie")
str_detect(x, "e")
```

```{r}
sum(str_detect(words, "^t"))
```

```{r}
mean(str_detect(words, "[aeiou]$"))
```

```{r}
no_vowels_1 <- !str_detect(words, "[aeiou]")
```

```{r}
no_vowels_2 <- str_detect(words, "^[^aeiou]+$")
identical(no_vowels_1, no_vowels_2)
```

```{r}
words[str_detect(words, "x$")]
```

```{r}
str_subset(words, "x$")
```

```{r}
df <- tibble(
  word = words, 
  i = seq_along(word)
)
df %>% 
  filter(str_detect(word, "x$"))
```

```{r}
x <- c("apple", "banana", "pear")
str_count(x, "a")
```

```{r}
mean(str_count(words, "[aeiou]"))
```

```{r}
df %>% 
  mutate(
    vowels = str_count(word, "[aeiou]"),
    consonants = str_count(word, "[^aeiou]")
  )
```

```{r}
str_count("abababa", "aba")
```

```{r}
str_view_all("abababababababa", "aba")
```

#### 14.4.1.1 Exercises

##### Find all words that start or end with x.
```{r}
head(words[str_detect(words, "^x")])
head(words[str_detect(words, "x$")])
```

##### Find all words that start with a vowel and end with a consonant.
```{r}
head(words[str_detect(words, "^[aeiou]")])
head(words[str_detect(words, "[^aeiou]$")])
```

### Extract matches

```{r}
length(sentences)
```

```{r}
head(sentences)
```

```{r}
colours <- c("red", "orange", "yellow", "green", "blue", "purple")
colour_match <- str_c(colours, collapse = "|")
colour_match
```

```{r}
has_colour <- str_subset(sentences, colour_match)
matches <- str_extract(has_colour, colour_match)
head(matches)
```

```{r}
more <- sentences[str_count(sentences, colour_match) > 1]
str_view_all(more, colour_match)
```

```{r}
str_extract(more, colour_match)
```

```{r}
str_extract_all(more, colour_match)
```

```{r}
str_extract_all(more, colour_match, simplify = TRUE)
```

```{r}
x <- c("a", "a b", "a b c")
str_extract_all(x, "[a-z]", simplify = TRUE)
```

#### 14.4.2.1 Exercises
```{r}
colors <- "(^red|orange|yellow|green|blue|purple)"
has_colors <- sentences %>%
  str_subset(colors) %>%
  head(10)
has_colors %>%
str_view_all(colors)
```


### Grouped matches

```{r}
noun <- "(a|the) ([^ ]+)"

has_noun <- sentences %>%
  str_subset(noun) %>%
  head(10)
has_noun %>% 
  str_extract(noun)
```

```{r}
has_noun %>% 
  str_match(noun)
```

```{r}
tibble(sentence = sentences) %>% 
  tidyr::extract(
    sentence, c("article", "noun"), "(a|the) ([^ ]+)", 
    remove = FALSE
  )
```

#### 14.4.3.1 Exercises
```{r}
number_word <- "(four|one|two|three|four|five|six|seven|eight|nine|ten) ([^ ]+)"

has_number_word <- sentences %>%
  str_subset(number_word) %>%
  head(701)
has_number_word %>% 
  str_extract(number_word)
```


### Replacing matches

```{r}
x <- c("apple", "pear", "banana")
str_replace(x, "[aeiou]", "-")
```

```{r}
str_replace_all(x, "[aeiou]", "-")
```

```{r}
x <- c("1 house", "2 cars", "3 people")
str_replace_all(x, c("1" = "one", "2" = "two", "3" = "three"))
```

```{r}
sentences %>% 
  str_replace("([^ ]+) ([^ ]+) ([^ ]+)", "\\1 \\3 \\2") %>% 
  head(5)
```

#### 14.4.4.1 Exercises

```{r}
x <- c("\\\\\\\\")
str_replace_all(x, c("\\\\" = "////"))
```


### Splitting

```{r}
sentences %>%
  head(5) %>% 
  str_split(" ")
```

```{r}
"a|b|c|d" %>% 
  str_split("\\|") %>% 
  .[[1]]
```

```{r}
sentences %>%
  head(5) %>% 
  str_split(" ", simplify = TRUE)
```

```{r}
fields <- c("Name: Hadley", "Country: NZ", "Age: 35")
fields %>% str_split(": ", n = 2, simplify = TRUE)
```

```{r}
x <- "First sentence, then wait! Another sentence here."
str_view_all(x, boundary("word"))
```

```{r}
str_split(x, " ")[[1]]
```

#### 14.4.5.1 Exercises

```{r}
x <- "apples, pears, and bananas"
str_view_all(x, boundary("word"))
```